\section{Introduction}
\label{section:introduction} 

\subsection{Motivation}

In the last few years we have seen a clear increase in the sale of mobile phones with an advanced OS, Internet access, GPS and large amounts of computing power. These so--called \emph{Smart Phones} are becoming increasingly common \justify{Find numbers or source}, replacing the standard mobile phone in the developed world. Developers are encouraged to write custom applications to these small mobile computers.
 
Many of these Smart Phones has a built-in GPS or WiFi triangulation tools, which can be access by the applications on the phone, making it possible to create services that uses the users location data --- such as directional services (for instance Google Maps) or social networking services (Foursquare).

In normal \anet research energy conservation is usually a high priority. Since \anet has mostly been used for research oriented sensor networks this makes sense. Sensor networks are usually deployed in inhospitable and inaccessible regions, which often makes it infeasible to recharge their power source, once it runs out of power. However, in the developed world recharging portable computers or Smart Phones is a trivial task. 

Mesh network could also play a vital role in humanitarian operations, such as after a earthquake, flood, hurricane etc. In such situations the infrastructure tends to be heavily damaged, and there will be a great need for coordination to maximise relief efforts. Imagine for instance that there has been an earthquake that has destroyed the wired infrastructure in a town. It is clear, that in order to the most good, the relief workers will need the ability to coordinate their efforts, otherwise some of the work is going to be duplicated, and things that needed immediate attention might be overlook for other, less urgent, problems. Since the local wired infrastructure has been destroyed, the relief workers cannot connect to the local Internet or telephone network. The relief workers may have brought in a \wbs (with some kind of long-lasting power source), which gives them access to the Internet. However, such a station is likely to be very expensive and bulky, and it is therefore infeasible to scatter these around the entire town. While it is almost a given that devices in its vicinity can connect to the Internet through WiFi, and the \wbs can improve the distance by boosting the signal, there is a upper limit to its range, before the noise-to-signal ratio becomes too great. \todo{think about whether I want to include refernce to \cite{disasterArea}}

An alternative could be to manually create a wired network, but this can be very time consuming, and prone to disruption as people or vehicles may destroy the connection. A third alternative may be to use wireless forwarding stations (WFS) -- these WFS must be placed such that they can both send and receive information from the \wbs and in this way \todo{Make a graph of this} make it possible to extend the area that can receive wireless Internet. However, these WFS are still an extra burden for the relief workers, and it is unlikely that they will be able to cover every knock and cranny. One way to remove the WFS' is to make each node make each computer node in the network propagate the wireless Internet signal. Preferably we would like to see this networks' topology to be made without human intervention, since this could prove rather onus. 

Fortunately, this is the area \anet, which for long has been studied in order to improve sensor data collection from inhospitable locations, like dense jungle or the arctic. \anet solutions include both \todo{include reference} a way to create the network topology, but also how to route it so that the network maximises its lifetime (since nodes may fall out if their power supply is exhausted). However, since we wanted this network to help relief workers, this network would be most useful if the relief workers could move around with smart-phones or laptops, all the while receiving and sending information. Doing this in a \anet is not practical, as they are optimized for a permanent network. Instead, we should try to use a \manet, which takes moving nodes as the baseline. Not surprisingly, designing \manet brings many new challenges that was not present in an \anet, since nodes can join and leave a \manet with an ease that is not possible in an \anet.  

Preferably we would like to have these \manet connected to the Internet, but this is not required for the \manet to work, and even without access to the Internet the \manet could still  be very useful for internal communication and file-sharing. 
 
\todo{finish this section and tighten it up}

\subsection{Expectations to the Reader}
I expect the reader to have a university level understanding of Computer Networking\footnote{For instance having read \cite{ComNet} or similar}, especially when it comes to understanding the building blocks of the network stack, as well as the most common problems with wireless networks. I also expect the reader to have a university level understanding of Algorithms\footnote{Having read \cite{algo} or similar}. \todo{Give this a few extra passes}

\subsection{Scope and Limitations}
While important in a real world application, I will in this thesis not concern myself with security, neither in the form of the authentication, anonymity of the sender or [something about rooting I think it might be possible]. For a treatment of this subject see \cite{trustedRouting}. Furthermore, if there does not exist a path between two nodes, I will not make any guarantee that the message will ever arrive at the sink. I will furthermore assume that all nodes have the same broadcast radius. \todo{finish this}

\hide{
\subsection{Node specifications}
It is clear that the properties of the mobile nodes in the system are going to influence the nature of the solution, I will therefore in this section define the specifications of the nodes I will work on in this project. The factors that I deem to be important are: 
\begin{description}
\item[Power:] How much does each node have? Can their supply be recharged?
\item[Mobility:] How mobile are the nodes? How fast can they move, and what can we expect their movement pattern to be? Can the nodes only realistically be integrated in a vehicles, or can a person be expected to carry them around on their person?
\item[(something - forgot what)]
\item[Amount of data to transmit\todo{Find better name than amount of data to transmit}:] 
\end{description}

As noted in section~\todo{find section where I mention power for the first time}, I would in this project like to look at power affluent devices, i.e. devices that either has access to large amounts of energy, or which has an adequate amount of energy, but which can easily be recharged. In this project I pick latter version, on the ground that such an arrangement is more realistic for a mobile device, and the setting it is going to be used in\footnote{It is unlikely that a Smart Phone will be able to transmit for 10 hours straight, or that the average user is going to transport a large capacity battery.}. In such a situation it is clear that my solution must be able to handle that nodes can suddenly disappear from the network from one moment to the next.

Concerning mobility I think the most interesting movement pattern would be people moving within streets or within buildings. I can justify this by the fact that 
}


\hide{
\subsection{Anatomy of the solution}

A real-life implementation of a Mobile Ad-Hoc Network will need the following elements

\begin{description}
\item[Entering and Merging networks:] When a node enters a new network, it needs to inform the network of its presence. Furthermore, when the node is already part of a secondary network (none of which are in the first network), then the networks need to be merged together.  
\item[Coordinate scheme:] Most routing schemes uses coordinates in order to find an efficient path, for any given metric, to a node. Therefore we must decide how the nodes are going to find their location. This is also a question about the hardware the nodes have (such as GPS or WiFi, and the required energy to power it for the nodes duration in the network), and some solutions   

Are the nodes going to use their position from a GPS system, or are they going to use some kind relative position.
\item[Node location mechanism:] There needs to be a mechanism so that a node can retrieve the coordinates of another nodes it wants to come into contact with. 
\item[Routing algorithm:] How are the nodes going to route. 
\item[Establishment of the underlying graph:]  Some routing algorithms will only work if the graph they are working on have certain properties\footnote{One such assumption might be that the graph is planar i.e. it can be drawn in the plane such that no edge crosses another}.
\item[Hierarchy:] Whether or not the nodes are going to route through levels of so-called \emph{cluster nodes} that will service numerous nodes.
\end{description} 

Furthermore, for the simulation we will have to decide upon a movement model (or find a number of real world traces we find representative) to test the effectiveness of solution. In this thesis I have decided to include \todo{decide upon which of these levels that I am going to implement, and which I will just assume have been solved}

\subsection{Goals}
In this section I will state the goals I will work toward in my this project:

\begin{description}
\item[Fairness:] Although I in this project assume that the devices in the Mesh network has plenty of power, I will still strive to ensure that the communication load across the network is somewhat fairly distributed. There are several reasons for this. Even if we assume that the devices that makes up the network can easily be recharged, extensive use of a node may deplete its current energy stores, and make it fall of the network. On the bandwidth side of the equation, it is likely a user will opt to retract his device from the network if he finds it is unable or slow to transmit even small messages (which might be the case if his device is heavily used by the network). Even with such a fairness system in place, unfair situations might still occur --- like when a single node becomes the bottleneck between 2 node clusters. 
\item[make more goals] \todo{add more goals or more this}
\end{description}

\subsection{Background Material Used}
For this project I have read the following articles for background information: ``A Survey on Wireless Mesh Networks'' \cite{martinWirelessSurvey}, ``Mesh networks: commodity multihop ad hoc networks''\cite{martinMeshnetwork} \todo{continue to include the articles you have read and are useful, but that have not been included directly}

\subsection{Source Code and Test Data}
The source code, report and test data can be found in my GitHub repository at \url{https://github.com/kjmikkel/MasterThesis} and is released under the \todo{add license the stuff is released under}.

\subsection{Overview}
In this section I give a quick overview of what the different sections in report will cover.
\begin{description}
\item[Section \ref{section:introduction}] The introduction to the report, which will introduce the problem of Mobile Ad-Hoc Networking.
\item[\todo{Add other sections}]
\end{description}
}
